---
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
library(tidyverse)
options(scipen = 999)

blocks.to.neighborhoods <- read_csv("Crosswalks/2010CensusBlocks_to_Neighborhoods.csv") %>%
  mutate(block = as.character(block))

median.age <- readRDS("SourceData/CensusBlocks2010_MedianAge.rds") %>%
  mutate(age = replace(age, age == 0, NA))
total.pop <- readRDS("SourceData/CensusBlocks2010_TotalPop.rds")
explore.accuracy <- readRDS("SourceData/AccuracyComparison.rds") %>%
  rename(`highest match` = group)
```

# Milwaukee-Geo-Crosswalks
This repo creates crosswalks between common Milwaukee geographies, e.g census tracts and neighborhood.

**Download the crosswalks as CSVs**

* [Census blocks (2010) to voting wards (2018)](https://github.com/jdjohn215/Milwaukee-Geo-Crosswalks/blob/master/Crosswalks/2010CensusBlocks_to_2018VotingWards.csv)
* [Census blocks (2010) to neighborhoods](https://github.com/jdjohn215/Milwaukee-Geo-Crosswalks/blob/master/Crosswalks/2010CensusBlocks_to_Neighborhoods.csv)
* [Census tracts (2017) to neighborhoods](https://github.com/jdjohn215/Milwaukee-Geo-Crosswalks/blob/master/Crosswalks/2017CensusTracts_to_Neighborhoods.csv)
* [Voting wards (2018) to neighborhoods](https://github.com/jdjohn215/Milwaukee-Geo-Crosswalks/blob/master/Crosswalks/2018VotingWards_to_Neighborhoods.csv)

## What is this?
When working with spatial data, the info we want is often available for a different kind of geography than the areas we care about. For example, some Census tracts are located entirely within neighborhoods, but others straddle two or more. The crosswalk files I've provided above allow data users to conveniently (dis)aggregate data from geography into another.

## Method
A quick way of creating a crosswalk is just to intersect the land areas of the two geographies. If 45% of geography A's land lies within geography B, then 45% of geography A's data is assigned to geography B. My method is a bit more sophisticated. I take the centroid of every residential parcel in the City of Milwaukee and intersect it with the data source geography. Then I calculate the proportion of *bedrooms* in geography A which are also in geography B.

## How it works
Here is what the crosswalks look like.

There is 1 row for every target geography TO data geography combination--in this case Census blocks to neighborhoods. The first line tells us that there are 154 bedrooms in both Alcott Park and Census block 550790195001005. These 154 bedrooms make up 100% of the bedrooms in the Census block, and they make up 9.4% of the bedrooms in Alcott Park.

```{r, echo=FALSE}
head(blocks.to.neighborhoods) %>%
  knitr::kable()
```

Here are the 2010 decennial census population counts for each block.
```{r, echo=FALSE}
head(total.pop) %>%
  knitr::kable()
```

Merge them. Calculate the population in each neighborhood/block combo by multiplying the pct.of.block adjustment factor by the population of the block. Group by neighborhood, then summarize.
```{r, warning=FALSE, message=FALSE}
blocks.to.neighborhoods %>%
  inner_join(total.pop) %>%
  mutate(pop.in.neighborhood = pop * pct.of.block) %>%
  group_by(neighborhood) %>%
  summarise(pop = sum(pop.in.neighborhood)) %>%
  arrange(desc(pop))
```

The above process works as long as you are dealing with a population count. If you want to (dis)aggregate a summary value (like median age), you can use the `pct.of.neighborhood` value to calculate a weighted mean.
```{r, warning=FALSE, message=FALSE}
blocks.to.neighborhoods %>%
  inner_join(median.age) %>%
  group_by(neighborhood) %>%
  summarise(median_age = weighted.mean(x = age, w = pct.of.neighborhood, na.rm = T)) %>%
  arrange(desc(median_age))
```

## How accurate is it?
The smaller the data source geography, the more likely that it will fit entirely within the target geography. The table below shows how each crosswalk performs.

* 99.4% of Census blocks fell entirely within a single neighborhood.
* 99.9% of Census blocks fell entirely within a single ward.
* Tracts are much more likely to straddle geographies. Just 30% fell within a single neighborhood. 10.5% of tracts were so divided that the largest tract/neighborhood combo still made up less than 50% of the tract.
* Wards can be merged to neighborhoods more accurately than tracts, but less accurately than blocks. 41% of wards lie entirely within a single neighborhood.

```{r, echo=FALSE}
explore.accuracy %>%
  knitr::kable(digits = 2)
```

